\bf An observational study of papers published in \emph{PLOS ONE} and studies posted to a trial registry.

### 2021-05-24 manuscript/asa_template/asa_template_almostdone.tex
update manuscript file and bib file
update tex files

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

Two statisticians on this paper (AB and NW) have heard researchers admit
that they have copied-and-pasted their statistical methods sections from
other papers, regardless of whether they are appropriate. The aim of
this paper is to use text-mining methods to analyse the content of
statistical methods sections included in common research outputs.
Results are then evaluated to estimate the extent that researchers are
using cut-and-paste or `boilerplate' statistical methods sections.
Boilerplate text is that ``which can be reused in new contexts or
applications without significant changes to the original''
\citep{Wikipedia}. Use of these methods sections indicates that little
thought has gone into the statistical analysis.

@@

Data included in our analysis were downloaded on 3 July 2020. For
records that did not pass Step 2, we reviewed where initial search terms
appeared in the full-text, to evaluate the likely proportion of
statistical methods sections that were missed.

@@

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The
total number of API search results peaked at volumes 8 (n = 19,045) and
9 (n = 19,045) in years 2013 and 2014. This trend aligned with the total
number of papers published in \emph{PLOS ONE} over the same period. The
percentage of records that included a statistical methods section by
volume based on our proposed matching criteria varied between 64\%
(volume 2) and 86\% (volume 9). All DOIs included Biology and life
sciences (n = 107,584), Earth sciences (n = 7,605) and/or Computer and
information sciences (n = 5,190) in their top 3 subject classifications
(Figure \ref{fig:plos-n}B).

@@

The topic clouds for ten clusters are in Figure \ref{fig:plos-clouds}.
Frequently occurring words reflected the use of statistical software
(topics 3 and 5), descriptive statistics (topic 6), group based
hypothesis testing (topics 1 and 4) and definitions of statistical
significance (topics 1 and 9). Statistical methods sections related to
regression (topic 2) and meta-analysis (topic 7) were also identified.
Examples of boilerplate text for selected topics are presented in Table
\ref{tab:plos-example-boilerplate}.

Topics related to statistical software differentiated between Prism
GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%).
Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
Diego, CA) was used for all analyses'' was found in 1,941 studies.
Similar phrasing was identified for topic 5: ``SPSS software version
17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'', with
similar text found in 441 studies.

@@

\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE datase (sentence level). Statistical methods text corresponds to the DOI with the highest topic coherence score within a topic}
\centering
\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
\hline
Topic & Statistical methods text & Matching DOIs\\
\hline
Topic 1 & Students t-test was used for statistical analysis. & 29\\
\cline{2-3}
 & A p value of $<$ 0.05 was considered statistically significant. & 1883\\
\hline
Topic 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
\hline
Topic 4 & Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
\hline
Topic 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
\hline
Topic 6 & All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
\hline
Topic 9 & Statistical significance was determined by Students t-tests. & 49\\
\cline{2-3}
& $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
\hline
\end{tabular}
\end{table}
\end{landscape}

Definitions of statistical significance featured strongly in topic 1 (n
= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%). Topic 1 reflected
applications of two-tailed and unpaired Student's t-tests, however,
instances of boilerplate text emphasised the 5\% level of statistical
significance (\ref{tab:plos-example-boilerplate}). In contrast, Topic 9
focused on multiple thresholds for declaring statistical significance by
asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
practice that has been criticised \citep{Wasserstein2019}.

Group-based hypothesis testing was a recurring theme across topics, with
text descriptions varying based on method(s) used. One-way analysis of
variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
with common methods for performing post-hoc multiple comparisons. Based
on the Jaccard index, 1 in 5 studies were matched to the sentence
``Significant differences were determined using analysis of variance
(ANOVA) follwo by Tukey post-hoc tests for multiple comparisons''.
Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
mentions of descriptive statistics for summarising continuous variables,
namely means with standard deviations or standard errors.

@@

In other cases, text had been slightly modified to account for changes
in primary and secondary outcomes. Examples of these text changes were
found in topic 2 (n = 1834) and topic 4 (n = 909); identified instances
related to sample size calculations for patient recruitment to different
studies (see Supplementary File 3).

Since studies included in this dataset outlined planned analyses, we
suspected that some studies did not specify statistical methods because
they had yet to consult with a statistician. Targeted searches for the
term ``statistician'' across all topics returned xx matching records,
with examples including ``A statistician employed by hospital was used''
and ``Pilot study at this point will use a statistician professionally
to determine sample size calculations as required''.

@@

\bibliographystyle{agsm}
\bibliography{references.bib}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

A potential contributor to poor statistical reporting is the temptation for researchers to replicate
descriptions of statistical methods, in an effort to make their papers resemble those of their peers 
and increase perceived chances of publication  \citep{Diong2018}. Two statisticians on this paper 
(AB and NW) have heard researchers admit
that they have copied-and-pasted their statistical methods sections from
other papers, regardless of whether they are appropriate. For this paper, we applied a text-based 
clustering method to analyse the content of statistical methods sections included in common scientific research outputs.
Clustering results are then evaluated to estimate the extent that researchers are
using cut-and-paste or `boilerplate' statistical methods sections.
Boilerplate text is that ``which can be reused in new contexts or
applications without significant changes to the original''
\citep{Wikipedia}. Use of these methods sections indicates that little
thought has gone into the statistical analysis.

@@

Data included in our analysis were downloaded on 3 July 2020. For
records that did not pass Step 2, we reviewed where initial search terms
appeared in the full-text, to determine the proportion of
statistical methods sections that were missed under our chosen approach.

@@

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The
total number of API search results peaked at volumes 8 (n = 19,045) and
9 (n = 19,045) in years 2013 and 2014. This trend aligned with the total
number of papers published in \emph{PLOS ONE} over the same period. The
percentage of records that included a statistical methods section varied between 64\%
(volume 2) and 86\% (volume 9). All DOIs included Biology and life
sciences (n = 107,584), Earth sciences (n = 7,605) and/or Computer and
information sciences (n = 5,190) in their top 3 subject classifications
(Figure \ref{fig:plos-n}B).

@@

The topic clouds for ten clusters are in Figure \ref{fig:plos-clouds}.
Frequently occurring words reflected the use of statistical software
(topics 3 and 5), descriptive statistics (topic 6), group based
hypothesis testing (topics 1 and 4) and definitions of statistical
significance (topics 1 and 9). Topics related to
regression (topic 2) and meta-analysis (topic 7) were also identified.
Examples of boilerplate text for selected topics are presented in Table
\ref{tab:plos-example-boilerplate}.

Topics related to statistical software differentiated between Prism
GraphPad (topic 3: n = 9974; 8.6\%) and SPSS (topic 5: n = 9648; 8.3\%).
Within topic 3, the sentence ``GraphPad Prism (Graphpad Software, San
Diego, CA) was used for all analyses'' was matched in 1,941 studies based on our definition of boilerplate text.
Similarly, for topic 5, ``SPSS software version
17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis'' returned matches for 441 studies.

@@

\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE datase (sentence level). Statistical methods text corresponds to the DOI with the highest topic coherence score within a topic}
\centering
\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
\hline
Topic & Statistical methods text & Matching DOIs\\
\hline
Topic 1 & $^{\dagger}$Students t-test was used for statistical analysis. & 29\\
\cline{2-3}
 & $^{\dagger}$A p value of $<$ 0.05 was considered statistically significant. & 1883\\
\hline
Topic 3 & $^{\dagger}$GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
\hline
Topic 4 & $^{\dagger}$Significant differences were determined using analysis of variance (ANOVA) followed by tukey post-hoc tests for multiple comparisons. & 2216\\
\hline
Topic 5 & $^{\dagger}$SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
\hline
Topic 6 & $^{\dagger}$All results are expressed as means $\pm$ standard deviation ($\pm$ SD). & 102\\
\hline
Topic 9 & $^{\dagger}$Statistical significance was determined by Students t-tests. & 49\\
\cline{2-3}
& $^{\dagger}$ $^*$p $<$ 0.05, $^{**}$p $<$ 0.01, $^{***}$p $<$ 0.001. & 6\\
\hline
\end{tabular}
\end{table}
\end{landscape}

Definitions of statistical significance featured strongly in topic 1 (n
= 3784; 3.2\%) and topic 9 (n = 6195; 5.3\%). Topic 1 reflected
applications of two-tailed and unpaired Student's t-tests, however,
instances of boilerplate text emphasised the 5\% level of statistical
significance (Table \ref{tab:plos-example-boilerplate}). In contrast, Topic 9
focused on multiple thresholds for declaring statistical significance by
asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
practice that has been criticised \citep{Wasserstein2019}.

Group-based hypothesis testing was a recurring theme across topics, with
text descriptions varying based on method(s) used. One-way analysis of
variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
with common methods for performing post-hoc multiple comparisons. Based
on the Jaccard index, 1 in 5 studies were matched to the sentence
``Significant differences were determined using analysis of variance
(ANOVA) followed by Tukey post-hoc tests for multiple comparisons''.
Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
mentions of descriptive statistics for summarising continuous variables,
namely means with standard deviations or standard errors.

@@

In other cases, text had been slightly modified to account for changes
in primary and secondary outcomes. Examples of these text changes were
found in topic 2 (n = 1834) and topic 4 (n = 909); identified instances
related to sample size calculations for patient recruitment to different
studies.

Since studies included in this dataset outlined planned analyses, we
hypothesised that some studies did not specify statistical methods because
they had yet to consult with a statistician. Targeted searches for the
term ``statistician'' across all topics returned xx matching records,
with examples including ``A statistician employed by hospital was used''
and ``Pilot study at this point will use a statistician professionally
to determine sample size calculations as required''.

@@

\bibliographystyle{agsm}
\bibliography{references}
